---
title: "Google Data Analytics Capstone Project"
author: "Christian Soenggoro"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the necessary libraries

```{r message=FALSE, warning=FALSE}
library(readr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(janitor)
library(ggplot2)
library(lubridate)
library(skimr)
library(forcats)
library(scales)
library(mapview)
library(here)
```

```{r message=FALSE}
here("202004-divvy-tripdata.csv")
```

### Load the data

```{r message=FALSE, warning=FALSE}
trips0 <- read_csv("Google Data Analytics/course 7 R code/202004-divvy-tripdata.csv")


```

### Check for NA values

```{r}
trips1 = trips0 %>% 
  is.na() %>% 
  colSums()
```

### Remove NA values and automatically clean column names

```{r}
trips2 = trips0 %>% 
  remove_empty(which = c("rows","cols")) %>% 
  clean_names()
```

### Change the latitude and longitude format from character to numeric type

```{r}
trips2 = trips2 %>% 
  mutate(start_lat = as.numeric(start_lat),
         start_lng = as.numeric(start_lng),
         end_lat = as.numeric(end_lat),
         end_lng = as.numeric(end_lng))
```

### Change format to date-time

```{r}
trips3 = trips2 %>% 
  mutate(started_at = as_datetime(started_at))
```

### Make new columns based on date-time information

```{r}
trips4 = trips3 %>% 
  mutate(hour_start = hour(started_at),
         week = week(started_at),
         weekday = wday(started_at, label =TRUE, abbr = FALSE),
         day = day(started_at),
         trip_time = difftime(ended_at, started_at,units = "min"))
```

### Change to factor data type and rename column for easier understanding

```{r}
trips5 = trips4 %>% 
  mutate(rideable_type = as_factor(rideable_type),
         member_casual = as_factor(member_casual)) %>% 
  rename(bikes = rideable_type,
         users = member_casual)
```

### Check for duplicate data

```{r}
trips5 %>% 
  duplicated() %>% 
  sum()
```

### Filter unlogical trip time

```{r message=FALSE, warning=FALSE}
trips6  = trips5 %>% 
  filter(between(trip_time,1,1440)) %>% 
  drop_na(end_lat, end_lng)
```

### Select columns that contains date-time information

```{r}
trips_time = trips6 %>% 
  select(ride_id,bikes,users,hour_start,week,weekday,day,trip_time)
```

### Select columns that contains location information

```{r}
trips_location =trips6 %>% 
  select(ride_id,bikes,users,start_station_name,start_lat,start_lng,end_station_name,start_station_id,
         end_lat,end_lng,end_station_id)
```

# Data Visualization

### Set plot theme

```{r}
newtheme <- theme_light() + 
  theme(plot.title = element_text(color = "#002949", face = 'bold', size =12),
        plot.subtitle = element_text(color = "#890000", size = 10),
        plot.caption = element_text(color = '#890000', face = 'italic', size =8),
        panel.border = element_rect(color = "#002949", size = 1),
        legend.position = "right",
        legend.text = element_text(colour="blue", size=10, face="bold"),
        legend.title = element_text(colour="blue", size=10, face="bold"),
        #legend.position='none',
        axis.title.x = element_text(colour = "#890000"),
        axis.title.y = element_text(colour = "#002949"),
        axis.text.x = element_text(angle = 45, hjust = 1, color = '#890000'),
        axis.text.y = element_text(angle = 45, hjust = 1, color = '#002949'),
        axis.line = element_line(color = "#002949", size =1),
  )

theme_set(newtheme)
```

## Hourly Data Visualization

```{r}
ride_hours = trips_time %>% 
  group_by(users,hour_start) %>% 
  summarise(nr_rides = n(),
            mean_time = mean(trip_time),
            total_time = sum(trip_time))
```

### Hourly number of rides

```{r}
ggplot(data = ride_hours,aes(x=hour_start, y =nr_rides)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Number of Trips per Hour" ,subtitle = "Number of Trips for every Hours segmented by Users",
       caption = "Figure 1" ,
       x= "hour of the day",
       y= "number of rides")+
  theme()
```

### Hourly Average trip time

```{r}
ggplot(data = ride_hours,aes(x=hour_start, y =mean_time)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Average duration of Trips per Hour" ,subtitle = "Average duration of Trips for every Hours and segmented Users",
       caption = "Figure 2" ,
       x= "hour of the day",
       y= "Average duration of rides")+
  theme()
```

### Hourly Total trip time

```{r}
ggplot(data = ride_hours,aes(x=hour_start, y =total_time)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Total duration of Trips per Hour" ,subtitle = "Total duration of Trips for every Hours and segmented by Users",
       caption = "Figure 3" ,
       x= "hour of the day",
       y= "Total time of rides")+
  theme()
```

## Day of the week visualization

```{r}
ride_weekly = trips_time %>% 
  group_by(users,weekday) %>% 
  summarise(nr_weekly = n(),
            mean_week = mean(trip_time),
            total_week = sum(trip_time))
```

### Day of the week number of rides

```{r}
ggplot(data = ride_weekly,aes(x=weekday, y = nr_weekly)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Number of Trips per Week" ,subtitle = "Number of Trips for day of the week and segmented by Users",
       caption = "Figure 4" ,
       x= "Week",
       y= "Number of rides")+
  theme()
```

### Day of the week Average trip time

```{r}
ggplot(data = ride_weekly,aes(x=weekday, y = mean_week)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Average duration of Trips per Week" ,subtitle = "Average duration of Trips for day of the week and segmented by Users",
       caption = "Figure 5" ,
       x= "Week",
       y= "Average duration of rides")+
  theme()
```

### Day of the week Total trip time

```{r}
ggplot(data = ride_weekly,aes(x=weekday, y = total_week)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Total duration of Trips per Week" ,subtitle = "Total duration of Trips for day of the week and segmented by Users",
       caption = "Figure 6" ,
       x= "Week",
       y= "Total duration of rides")+
  theme()
```

## Day of the month Visualization

```{r}
ride_day = trips_time %>% 
  group_by(users,day) %>% 
  summarise(nr_day = n(),
            mean_day = mean(trip_time),
            total_day = sum(trip_time))
```

### Day of the month number of rides

```{r}
ggplot(data = ride_day,aes(x = day, y = nr_day)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Number of Trips per day of the month" ,subtitle = "Number of Trips for day of the month and segmented by Users",
       caption = "Figure 7" ,
       x= "day",
       y= "Number of rides")+
  theme()
```

### Day of the month Average trip time

```{r}
ggplot(data = ride_day,aes(x=day, y = mean_day)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Average duration of Trips per day of the month" ,subtitle = "Average duration of Trips for day of the month and segmented by Users",
       caption = "Figure 8" ,
       x= "day",
       y= "Average duration of rides")+
  theme()
```

### Day of the month Total trip time

```{r}
ggplot(data = ride_day,aes(x=day, y = total_day)) +
  geom_bar(aes(fill=users),position = "dodge",stat = "identity")+
  scale_y_continuous()+
  labs(title ="Total duration of Trips per day of the month" ,subtitle = "Total duration of Trips for day of the month and segmented by Users",
       caption = "Figure 9" ,
       x= "day",
       y= "Total duration of rides")+
  theme()
```

## Visualize by location

```{r}
pop_start_station = trips_location %>% 
  group_by(
    users, start_station_name, start_lat, start_lng
  ) %>% 
  summarise(
    nr_rides_start = n()
  ) %>% 
  arrange(-nr_rides_start)


```

```{r}
pop_end_station = trips_location %>% 
  group_by(
    users, end_station_name, end_lat, end_lng
  ) %>% 
  summarise(
    nr_rides_end = n()
  ) %>% 
  arrange(-nr_rides_end)
```

### Top 10 Start station

```{r}
pop_start_station[1:10, ] %>% 
  ggplot(aes(start_station_name, nr_rides_start, fill = users))+
  geom_col(position = "dodge")+
  coord_flip()+
  labs(
    title = "Most Popular Start Stations",
    subtitle = "Top 10 most popular start stations",
    caption = "Fig 10 ",
    x = "station name",
    y = "number of trips"
  )+
  theme()
```

### Top 10 end station

```{r}
pop_end_station[1:10,] %>% 
  ggplot(aes(end_station_name, nr_rides_end, fill = users))+
  geom_col(position = "dodge")+
  coord_flip()+
  labs(
    title = "Most Popular End Stations Segmented by Users",
    subtitle = "Top 10 most popular end stations",
    caption = "Fig 11",
    x = "station name",
    y = "number of trips"
  )+
  theme()
```

### Mapview of 30 most popular start station

```{r}
pop_start_station[1:30, ] %>%
  mapview(
    xcol = "start_lng", 
    ycol = "start_lat",
    cex = "nr_rides_start",
    alpha = 0.9, 
    crs = 4269,
    color = "#8b0000",
    grid = F, 
    legend = T,
    layer.name = "30 Most Popular Start Stations")
```

### Mapview of 30 most popular end station

```{r}
pop_end_station[1:30,] %>% 
  mapview(
    xcol = "end_lng",
    ycol = "end_lat",
    cex = "nr_rides_end", # size of circle based on value size
    alpha = 0.9,
    crs = 4269,
    color = "#8b0000",
    grid = F,
    legend = T,
    layer.name = "30 Most Popular End Stations")
```
